Python as a machine learning language: experiences from the scikit-learn project

### 2013-04-27 api.tex

Construction and application of an estimator object
proceeds in a way that is somewhat reminiscent of partial function application.
The constructor, \texttt{\_\_init\_\_}\footnote{
  Strictly speaking, \texttt{\_\_init\_\_} is Python's
  class \textit{initialization} method,
  while \texttt{\_\_new\_\_} is the constructor.
  The distinction is not relevant to the present description of scikit-learn.}
takes a set of named (hyper-)parameters
and attaches these to the estimator as public attributes.\footnote{
  The Python language does not actually distinguish between public and private
  attributes and methods,
  but by convention, any identifier prefixed with an underscore (\texttt{\_})
  is considered ``a non-public part of the API\ldots
  an implementation detail and subject to change without notice.''
  % TODO cite Python tutorial here
  }
It never performs actual learning, nor does it even see a training set.

@@

% TODO also discuss scipy.sparse
Input to an estimator is typically given as NumPy arrays \citep{vanderwalt2011};
typically, a clustering or dimensionality reduction algorithm
takes a single NumPy array, conventionally called $X$ of size $n \times d$
to represent $n$ samples of $d$ features each.
A regression or classification estimator takes, at training time,
an additional array of size $n$, by convention called $y$,
that represents the target variable.
While the NumPy array may seem like a ``bare'' representation of data
when compared to more object-oriented representations
(again, Weka is an example of this style of API),
it brings the prime advantages of allowing \textit{vector operations}
on entire arrays,
which are often orders of magnitude faster than the corresponding Python loops
while at the same time keeping the code short and readable.
Although conversion to the NumPy format may seem like a burden on the user,
we argue that this is not the case.
Python's lists (its dynamic array structure)
are easily transformed into NumPy arrays
and many scientific users of Python will already be handling such arrays,
since they are pervasive in other science-oriented Python packages.
For tasks where input is likely to consists of text files
or semi-structured objects, ``vectorizers'' are provided
that efficiently convert from such data formats to the NumPy format.
Finally, offering a custom datatype for samples
would require a conversion \textit{in any case}.

>>>>

Construction and application of an estimator object
proceeds in a way that is somewhat reminiscent of partial function application.
The constructor, \texttt{\_\_init\_\_}\footnote{
  Strictly speaking, \texttt{\_\_init\_\_} is Python's
  class \textit{initialization} method,
  while \texttt{\_\_new\_\_} is the constructor.
  The distinction is not relevant to the present description of scikit-learn.}
takes a set of named (hyper-)parameters
and attaches these to the estimator as public attributes.\footnote{
  The Python language does not actually distinguish between public and private
  attributes and methods,
  but by convention, any identifier prefixed with an underscore (\texttt{\_})
  is considered ``a non-public part of the API\ldots
  an implementation detail and subject to change without notice.''
  \Citep{pythontut}
  }
It never performs actual learning, nor does it even see a training set.

@@

% TODO also discuss scipy.sparse
Input to an estimator is typically given as NumPy arrays
\Citep{vanderwalt2011};
typically, a clustering or dimensionality reduction algorithm
takes a single NumPy array, conventionally called $X$ of size $n \times d$
to represent $n$ samples of $d$ features each.
A regression or classification estimator takes, at training time,
an additional array of size $n$, by convention called $y$,
that represents the target variable.
While the NumPy array may seem like a ``bare'' representation of data
when compared to more object-oriented representations
(again, Weka is an example of this style of API),
it brings the prime advantages of allowing \textit{vector operations}
on entire arrays,
which are often orders of magnitude faster than the corresponding Python loops
while at the same time keeping the code short and readable.
Although conversion to the NumPy format may seem like a burden on the user,
we argue that this is not the case.
Python's lists (its dynamic array structure)
are easily transformed into NumPy arrays
and many scientific users of Python will already be handling such arrays,
since they are pervasive in other science-oriented Python packages.
For tasks where input is likely to consists of text files
or semi-structured objects, ``vectorizers'' are provided
that efficiently convert from such data formats to the NumPy format.
Finally, offering a custom datatype for samples
would require a conversion \textit{in any case}.
